---
title: "30_machine_learning_model"
author: "Daniel Molnar"
date: "2023-06-02"
output:
  rmdformats::robobook:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Set Working Directory
if (rstudioapi::isAvailable()) {
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
}

# Install Libraries
# install.packages("zoo")

# Load Libraries
library(data.table)
library(plotly)
library(zoo)
library(caret)

# Set Seed
set.seed(123)
```

# Task: Predict the total_power that will be drained from the wallboxes on the next day using data of the last 14 days.

## Import Data

The **data is already preprocessed.** I just removed one additional feature we used for plotting in another task.

```{r import data}
wallboxes_jan_aug_DT <- fread("./data/preprocessed/total_power_jan-aug.csv")

wallboxes_jan_aug_DT[, battery_SOC := NULL]
```

## Feature Engineering

For this task I performed some feature engineering where I tried to enrich the data I have so far. I started off with two columns, namely *Date* and *total_power*. Finally, I came up with **4 new features:**

- **total_power_previous_day:** As the name already suggests I added a column containing the *total_power* that was used on the previous day.
- **total_power_same_day_previous_week:** Again very simple I added a column containing the *total_power* that was used on the same day last week.
- **average_total_power_last_seven_days:** For this I calculated the average *total_power* that was used during the last seven days.
- **is_weekend:** This column is set to 1 if the corresponding *Date* was a weekend (saturday, sunday; in my case "Samstag", "Sonntag" because of german language) or 0 if it was a normal weekday.

```{r feature engineering, warning=FALSE}
# 1. total_power used on previous day
wallboxes_jan_aug_DT[, total_power_previous_day := shift(total_power, 1)]

# 2. total_power used on the same day last week
wallboxes_jan_aug_DT[, total_power_same_day_previous_week := shift(total_power, 7)]

# 3. average total_power that was used in the last 7 days (rolling average)
wallboxes_jan_aug_DT <- wallboxes_jan_aug_DT %>%
                          mutate(average_total_power_last_seven_days = rollmean(total_power, k = 7, fill = NA, align = "right"))
                                 
# I found out that this calculation sums the last X values and divides it by X and then writes it to row X, this doesn't
# make sense because when we want to predict the total_power that will be used on a day then we of course don't know the
# total_power of this day already. Therefore all of the solutions have to be shifted by 1. So that for example the rolling average
# of the first 7 days is written in row 8. Meaning we can use this knowledge to predict total_power on day 8.

wallboxes_jan_aug_DT[, average_total_power_last_seven_days := shift(average_total_power_last_seven_days, 1)]

# 4. weekend or not
wallboxes_jan_aug_DT[, is_weekend := weekdays(Date)]
wallboxes_jan_aug_DT$is_weekend <- ifelse(wallboxes_jan_aug_DT$is_weekend %in% c("Samstag", "Sonntag"), 1, 0)
```

Let's have a look at the data so far.

```{r look at data}
str(wallboxes_jan_aug_DT)
summary(wallboxes_jan_aug_DT)
head(wallboxes_jan_aug_DT, n = 15)
```

As you can see I now have some NA values because I shifted the data around. Let's remove observations containing NAs in one of their columns.

```{r remove NA}
# colSums(is.na(wallboxes_jan_aug_DT))
wallboxes_jan_aug_DT <- na.omit(wallboxes_jan_aug_DT)
# str(wallboxes_jan_aug_DT)
# summary(wallboxes_jan_aug_DT)
head(wallboxes_jan_aug_DT, n = 15)
```

## Random Forest Models

Finally let's **train multiple random forest models and always try predicting the total_power that will be used on the next day using the data of the last 14 days.** After that I calculate the **average Mean Absolute Error (MAE)** of all models and the **average Mean Absolute Percentage Error (MAPE)** of all models so that I have some metrics I can use for evaluating the models.

```{r random forest without Date, warning=FALSE}
MAPE <- function(predicted, actual){
  mape <- mean(abs((actual - predicted)/actual))*100
  return (mape)
}

actuals <- wallboxes_jan_aug_DT
start_date <- as.Date("2022-01-08")
end_date <- as.Date("2022-01-22")
predicted_date <- list() # the date I tried to predict will get stored in this list
predictions <- list() # the predictions will get stored in this list
real_values <- list() # the actual values will get stored in this list
errors <- list() # the MAE for each model will get stored in this list
mape_errors <- list() # the MAPE for each model will get stored in this list

i <- 1
while (end_date <= as.Date("2022-08-21")) {
  
  # print(paste('Currently calculating prediction for', end_date))
  
  # use only the last 14 days to train a RF model
  training <- actuals[Date >= start_date & Date < end_date,]
  test <- actuals[actuals$Date == end_date]
  
  rf_model <- train(total_power ~ . - Date, 
                    data = training, 
                    method = "ranger")
  
  pred <- predict(rf_model, newdata = test) # predict 1 day
  predictions[[i]] <- pred # store the prediction
  predicted_date[[i]] <- end_date
  actual <- test$total_power
  real_values[[i]] <- actual
  errors[[i]] <- abs(actual - pred) # store the difference between actual and predicted value
  mape_errors[[i]] <- MAPE(pred, actual)
  
  i <- i+1
  
  start_date <- start_date + 1
  end_date <- end_date + 1
}

mean_absolute_error <- mean(unlist(errors))
mean_absolute_error

mean_absolute_percentage_error <- mean(unlist(mape_errors))
mean_absolute_percentage_error
```

Let's also look at a plot to better **visualize the results.** The first plot uses *lines+markers* to better identify the single points, the second one only uses *lines*.

```{r plots without Date}
# lines+markers
plot_ly() %>%
  add_trace(x = ~predicted_date, y = ~real_values, name = 'Actuals', type = 'scatter', mode = 'lines+markers') %>%
  add_trace(x = ~predicted_date, y = ~predictions, name = 'Predictions', type = 'scatter', mode = 'lines+markers') %>%
  layout(title = 'RF predictions using the last 14 days', xaxis = list(title="Date"), yaxis =list(title="kWh"))

# only lines
plot_ly() %>%
  add_trace(x = ~predicted_date, y = ~real_values, name = 'Actuals', type = 'scatter', mode = 'lines') %>%
  add_trace(x = ~predicted_date, y = ~predictions, name = 'Predictions', type = 'scatter', mode = 'lines') %>%
  layout(title = 'RF predictions using the last 14 days', xaxis = list(title="Date"), yaxis =list(title="kWh"))
```






